### 2023-04-25 thesis/MSECE_thesis.tex
Update bibtex references; update lit review, data sources.

\settowidth{\parindent}{}

@@

    \section{Abstract}
    \section{Introduction}
    \section{Literature Review}
    \section{Methodology}
        \subsection{Data Source(s)}
        \subsection{Data Analysis}
        \subsection{Preprocessing}
    \section{Results}
    \section{Discussion}
    \section{Conclusion}

>>>>

% \settowidth{\parindent}{}

@@

\section{Abstract}
\section{Introduction}
\section{Literature Review}
% from proposal; revise and add to this as more references are used
% transformer history
The transformer neural network architecture was proposed in 2017 for Neural Machine Translation tasks and immediately achieved
state-of-the-art (28.4 BLEU on WMT 2014 English-to-German; 41.8 BLEU on WMT English-to-French datasets)
\cite{vaswani_attention_2017}. Transformer architectures have been found to be extremely effective at learning representations
and understandings of languages to predict token probabilities as opposed to transforming one language into another
\cite{devlin_bert_2019,liu_roberta_2019}. Transformers have also been found to be very effective at
other NLP-related tasks such as prompt completion and sentiment analysis among others (i.e.~auto-regressive and sequence
classification tasks, respectively) \cite{lewis_bart_2019,radford_improving_2018}.

% ASR context (suggest some need for semi-supervised approaches and therefore LMs)
End-to-end generative models for automatic speech recognition models have made significant progress in recent years with
current state-of-the-art models achieving WERs as low as 2\% on LibriSpeech test sets
\cite{han_contextnet_2020,kriman_quartznet_2020,baevski_wav2vec_2020,li_jasper_2019}. This has led to the development of ASR
models for the aviation domain, specifically, in air traffic control communications
\cite{badrinath_automatic_2022,smidl_air_2019,zuluaga-gomez_automatic_2020,srinivasamurthy_semi-supervised_2017}.
However, due to the lack of transcribed data in the aviation domain
\cite{zuluaga-gomez_automatic_2020,srinivasamurthy_semi-supervised_2017,badrinath_automatic_2022,smidl_air_2019}
ASR models maintain relatively high WERs compared to their counterparts in the more generalized ASR domain
\cite{zuluaga-gomez_automatic_2020,badrinath_automatic_2022}. Transfer learning has even yielded limited results
in this domain (depending on model architecture and dataset quality)
\cite{badrinath_automatic_2022,zuluaga-gomez_automatic_2020}.

% Unsupervised/semi-supervised in general and LMs commonly used with them
Unsupervised and semi-supervised methodologies have become popular recently in attempts to address limited data availability
and develop new approaches towards modeling human speech (notably, wav2vec has achieved state-of-the-art performance with
very little training data)
\cite{baevski_wav2vec_2020,badrinath_automatic_2022,srinivasamurthy_semi-supervised_2017,zuluaga-gomez_contextual_2021}.
Language models are an integral part semi-supervised learning. They are used to obtain a certainty score (or uncertainty
score, as the case may be) for the predicted text, these are usually either word lattices or N-gram models
\cite{badrinath_automatic_2022,srinivasamurthy_semi-supervised_2017,zuluaga-gomez_contextual_2021}.
While these have proven to be adequate for most self-supervised learning tasks, they are hardly state-of-the-art.
Wav2vec, the current state-of-the-art unsupervised model (and possibly top performing overall), uses a contrastive process
between convolutional feature extraction and transformer context representations \cite{baevski_wav2vec_2020}.

% LMs and NLP in aero. domain
Various natural language processing methods have been applied to the aviation domain to help deal with miscommunications
and try to mitigate safety incidents \cite{ragnarsdottir_language_2003,tanguy_natural_2016,madeira_machine_2021}.
Some machine learning approaches have been implemented to analyze the text in aviation incident and safety reports to predict
contributing factors and topic models \cite{tanguy_natural_2016,madeira_machine_2021}. An ASR system with NLP post-processing
has also been proposed to analyze Air-Traffic communications and condense significant features (e.g.~weather, runway, and
heading info) into an XML language structure \cite{ragnarsdottir_language_2003}. Transformer language models such as BERT,
RoBERTa, and DeBERTa have been applied to notice to airmen (NOTAM) messages to perform named entity recognition (NER) tasks,
translation tasks (between notations; e.g.~NOTAM notation to Airlang to make parsing tasks easier) and reduce the workload
for pilots during long-haul flights \cite{arnold_knowledge_2022}. Transformer models have also been used for speaker role
identification tasks in the aviation domain (e.g.~identifying pilot versus controller in communications), specifically,
a pretrained BERT model was used and fine-tuned on problem specific data and compared to other models that performed
well at speaker and role identification tasks in general \cite{guo_comparative_2022}.
\section{Methodology}
\subsection{Data Source(s)}
There are four datasets used in this work. All are described in detail below. All data samples are text-based and are mined
from transcripts intended for Automatic Speech Recognition (ASR) tasks.

\textbf{Air Traffic Control Complete}. Hereafter this dataset will be referred to as ATCC. This data was collected from three major
airports in the United States, specifically, Boston Logan, Washington National, and Dallas-Fort Worth airports. The data is intended
to be used primarily for Automatic Speech Recognition tasks, therefore the data in this work was mined from the transcripts of those
data samples \cite{godfrey_air_1994}.

\textbf{Air Traffic Control 2}. This dataset is referred to as ATCO2 throughout the rest of this work. This data was collected as part
of an open source initiative to collect and label audio data from airports around the world. The data was originally used for
acoustic-based NLP tasks, such as ASR, in the aviation/aerospace domain(s). Similarly to ATCC, the data used in this work from ATCO2
is mined from the transcripts of the samples in the dataset \cite{szoke_detecting_2021}.

\textbf{Air Traffic Control Simulation}. This dataset is referred to as ATCOSIM throughout the rest of this thesis. The data in this
corpus was collected at an air-traffic control training and simulation center used for teaching and evaluating air-traffic control
and air-traffic management concepts. As with the other datasets in this work, this dataset was created for acoustic and audio-based
analysis and machine learning tasks such as ASR among others. Likewise, the data used in this work is mined from the transcripts
of the samples in the dataset \cite{hofbauer_atcosim_2008}.

\textbf{Czech ATC Corpus}. This dataset is referred to throughout the rest of this work as the ZCUATC corpus/dataset. The data in this
corpus was recorded at an airport in the Czech Republic and subsequently transcribed. This dataset differs the most from the other
datasets used in this work due to the location in which it was collected and transcribed, however a majority of the samples, if not all
of them, should conform the the ICAO standards of aviation communications, which is why it remains in this work. As with the other
audio-based corpora, the data used in this thesis was derived from the transcripts that correspond to the audio data in the ZCUATC
corpus \cite{smidl_air_2019}.

\subsection{Data Analysis}
% sequence lengths, unique tokens, total number of tokens, total number of samples, breakdown by each dataset
\subsection{Preprocessing}
% explain and justify what is removed from the text, how, and why
% present and explain any relevant algorithms where necessary
\subsection{Tokenization}
% present and explain the tokenization algorithms used (possibly which models they correspond to)
\section{Discussion}
\section{Results}
\section{Conclusion}
